---
title: "analysis_script.Rmd"
author: "DS421 Team"
date: "March 9, 2020"
output: html_document
---
Install Packages
```{r}
library(tidyverse)
source('scripts/adams_theme.R') #for ggplot2
```


Import data
```{r}
df <- read_csv('data/FireDataKlamath1987_part1.csv', col_types = cols(clusterID = col_factor())) %>%
  rbind(read_csv('data/FireDataKlamath1987_part2.csv', col_types = cols(clusterID = col_factor())))

eco_type_codes <- read_csv('data/landfire_ecotype_codes.csv')

```

clean data
```{r}
df <- df %>%
  select(-.geo) %>%
  #filter(FIRE_YEAR == 1994) %>%
  rename(pixelID = 'system:index')

```


Identifying patches with more than eight pixels
```{r}
clusters_with_more_than8_pixels <- df %>% group_by(clusterID) %>%
  summarise(Npixels = length(unique(pixelID))) %>%
  filter(Npixels > 8) %>%
  pull(clusterID)
```


Creating a landcover data frame
```{r}

landcover <- names(df)[grep(names(df), pattern = "NLCD")]

df_lc <- df %>%
  select(pixelID, clusterID, landcover) %>%
  gather(landcover, key = "NLCD_year", value = "landcover_code") %>%
  mutate(land_cover_name = case_when(
    (landcover_code == 71) ~ "grass_herb",
    (landcover_code == 51) ~ "shrub",
    (landcover_code == 42) ~ "evergreen_forest",
    (landcover_code == 43) ~ "mixed forest",
    (landcover_code == 31) ~ "barren_land",
    (landcover_code == 23) ~ "developed",
    (landcover_code == 41) ~ "deciduous_forest",
    (landcover_code == 11) ~ "open_water",
    (landcover_code == 81) ~ "pasture_hay",
    (landcover_code == 33) ~ "transitional_barren",
    (landcover_code == 52) ~ "shrub",
    (landcover_code == 21) ~ "developed",
    (landcover_code == 95) ~ "emergent_herb_wetlands",
    (landcover_code == 90) ~ "woody_wetlands",
    (landcover_code == 22) ~ "developed"
  )) %>%
  mutate(year = as.numeric(gsub("NLCD", "", NLCD_year))) %>%
  arrange(clusterID)

df_lc
```


Creating a tree canopy cover df
```{r}

TCC <- names(df)[grep(names(df), pattern = "TCC")]

df_tcc <- df %>%
  select(pixelID, clusterID, TCC) %>%
  gather(TCC, key = "TCCyear", value = "pct_TCC") %>% 
  mutate(year = as.numeric(gsub("TCC", "", TCCyear)))

df_tcc
```


Creating an NBR dataframe
```{r}
years <- as.character(1985:2017)
df_nbr <- df %>%
  select(pixelID, clusterID, years) %>%
  gather(years, key = "year_string", value = "NBR") %>%
  mutate(year = as.numeric(year_string)) %>%
  select(-year_string) %>%
  arrange(clusterID)

df_nbr
```


Creating a FACTS df
```{r}

facts_cols <- names(df)[grep(names(df), pattern = "facts")]
df_facts <- df %>%
  select(pixelID, clusterID, facts_cols) %>%
  gather(facts_cols, key = "facts_year", value = "factsTreatmentCode") %>%
  mutate(factsTreatment = case_when(
    (factsTreatmentCode == 0) ~ "none",
    (factsTreatmentCode == 4431) ~ "plant_trees",
    (factsTreatmentCode == 4432) ~ "replant_or_fill"
  )) %>%
  mutate(year = as.numeric(gsub("facts", "", facts_year))) %>%
  arrange(clusterID)

df_facts
```


Merging dfs together
```{r}
#mering dfs together
df <- df %>% select(pixelID, clusterID, FIRE_YEAR, FireID, burnSev, distance, northness, eastness, elevation, slope, b1, WID) %>%
  rename(ecotype_code = b1) %>%
  left_join(df_nbr, by = c("pixelID","clusterID")) %>%
  left_join(df_lc,c("pixelID","clusterID","year")) %>%
  left_join(df_facts,c("pixelID","clusterID","year")) %>% 
  mutate(burnSevLong = case_when(
    (burnSev == 0) ~ "background/no_data",
    (burnSev == 1) ~ "verylow/no_burn",
    (burnSev == 2) ~ "low",
    (burnSev == 3) ~ "med",
    (burnSev == 4) ~ "high"
  )) %>%
  mutate(eco_type = case_when(
    (ecotype_code == 549) ~ "MC_LM_Serp",
    (ecotype_code == 550) ~ "MC_UM_Serp",
    (ecotype_code == 551) ~ "MC_DrytoMes",
    (ecotype_code == 552) ~ "MC_Mes",
    (ecotype_code == 553) ~ "MixedOak",
    (ecotype_code == 554) ~ "MC_BO_LM",
    (ecotype_code == 555) ~ "Jeffrey_Ponderosa",
    (ecotype_code == 556) ~ "RedFir",
    (ecotype_code == 557) ~ "SubalpineWoodland",
    (ecotype_code == 558) ~ "Maritime_DFir_WHem"
  )) %>%
  filter(ecotype_code %in% eco_type_codes$eco_type_code) 

df
#adding the ecotype descriptions
```



Checking representation of forest types. Make this a table.
```{r}
#representation of forest types
df %>%
  group_by(eco_type) %>%
  summarise(number_pixels = length(unique(pixelID))) %>%
  drop_na() %>%
  mutate(sample_area_hectares = number_pixels * 900 / 10000) %>%
  ggplot(aes(eco_type, number_pixels)) +
  geom_bar(stat = "identity") +
  adams_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(label = "N pixels")



total_sample_area <- length(unique(df$pixelID)) * 900 / 10e4 
total_sample_area 

total_number_patches <- length(unique(df$clusterID))
total_number_patches
```


Exploring patch size in hectares and pixels
```{r}
df %>%
  group_by(clusterID) %>%
  summarise(num_pixels_per_cluster = length(unique(pixelID))) %>%
  mutate(cluster_area = num_pixels_per_cluster * 900 / 10000) %>%
  pull(cluster_area) %>% summary()
  
df %>%
  group_by(clusterID) %>%
  summarise(num_pixels_per_cluster = length(unique(pixelID))) %>%
  mutate(cluster_area = num_pixels_per_cluster * 900 / 10000) %>%
  ggplot(aes(num_pixels_per_cluster)) +
  geom_histogram() +
  xlim(c(0,100)) +
  adams_theme

length(unique(df$pixelID))
length(unique(df$pixelID)) * 900 / 10e4
  
```



Quantifying how many pixels went from non-forest back to forest in their post-fire regeneration trajectory
```{r}
`%notin%` <- Negate(`%in%`)

nonForest2001 <- df %>%
  filter(year == 2001, landcover_code %notin% 41:43) %>%
  pull(pixelID) %>% unique()

#how many pixels went from shrub to tree
df %>%
  filter(pixelID %in% nonForest2001) %>%
  filter(year == 2016, landcover_code %in% 41:43) %>%
  pull(pixelID) %>% unique() %>% length()



```




Adding the percent NBR recovey to the df.
%NBR recovery is the fraction of mean pre-fire NBR
```{r}
#creating pre-fire NBR values for each pixel
pre_fire_nbr <- df %>% 
  filter(year %in% 1985:1986) %>%
  group_by(pixelID) %>%
  summarise(preFireNBR = mean(NBR))

#adding the percent NBR recovery to the df
df <- df %>%
  left_join(pre_fire_nbr) %>%
  mutate(pct_nbr_recovery = NBR / preFireNBR)
```


Showing the distribution of NBR values for pixel classified as forest for NLCD
```{r}
# df %>%
#   drop_na(land_cover_name) %>%
#   filter(landcover_code %in% 41:43) %>%
#   ggplot(aes(NBR)) +
#   geom_histogram()


df %>%
  drop_na(land_cover_name) %>%
  filter(landcover_code %in% c(41:42, 71,51,52)) %>%
  ggplot(aes(x=NBR, fill=land_cover_name)) +
  geom_density( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("light green", "dark green", "yellow", "brown")) +
  #theme_ipsum() +
  labs(fill="") +
  adams_theme


```


plot patch-level NBR through time for a random sample of patches
```{r}
#patches <- sample(unique(df$clusterID),100)

df %>%
  group_by(FireID, clusterID, year) %>%
  summarise_if(is.numeric,mean) %>%
  #filter(clusterID %in% patches) %>%
  ggplot(aes(year,NBR,color = clusterID)) +
  geom_line() +
  adams_theme +
  theme(legend.position = "none")
```


Calculating the percent NBR recovery for each patch
```{r}
#patches <- sample(unique(df$clusterID),10)
df %>%
  group_by(clusterID, year) %>%
  #filter(WID > 0) %>%
  summarise_if(is.numeric,mean) %>%
  #filter(clusterID %in% patches) %>%
  ggplot(aes(year,pct_nbr_recovery,color = clusterID)) +
  geom_line() +
  ylim(c(-1,1.5)) +
  adams_theme +
  theme(legend.position = "none")
```



Histogram of percent recovery in 2000 (6 years after fire)
```{r}
df %>%
  group_by(pixelID, year) %>%
  summarise_if(is.numeric,mean) %>%
  filter(year == 1992) %>%
  ggplot(aes(x = pct_nbr_recovery, fill = factor(burnSev))) +
  geom_density(alpha=0.6, position = 'identity') +
  scale_fill_brewer(palette = "Reds") +
  xlim(c(-0.2,2)) +
  xlab(label = "% NBR Recovery 5 years post-fire") +
  adams_theme +
  labs(fill="Burn Severity") 
  
```

Density plot of percent recovery in 2017 (23 years after fire)
```{r}
df %>%
  group_by(pixelID, year) %>%
  summarise_if(is.numeric,mean) %>%
  filter(year == 2017) %>%
  ggplot(aes(x = pct_nbr_recovery, fill = factor(burnSev))) +
  geom_density(alpha=0.6, position = 'identity') +
  scale_fill_brewer(palette = "Reds") +
  xlim(c(-0.2,2)) +
  xlab(label = "% NBR Recovery 30 years post-fire") +
  adams_theme +
  labs(fill="Burn Severity") 
```


Creating variables showing the percent NBR recovery in 92,97,2017
```{r}
df <- df %>%
  filter(pct_nbr_recovery >= -10)

pctNBRrec2017 <- df %>%
  filter(year == 2017) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_2017 = pct_nbr_recovery)

pctNBRrec1997 <- df %>%
  filter(year == 1997) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_1997 = pct_nbr_recovery)

pctNBRrec1992 <- df %>%
  filter(year == 1992) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_1992 = pct_nbr_recovery)

df <- df %>%
  left_join(pctNBRrec2017) %>%
  left_join(pctNBRrec1997) %>%
  left_join(pctNBRrec1992)
```



Showing the perecent NBR recovery by burn severity class (10 years after fire)
```{r}
# df %>%
#   ggplot(aes(factor(burnSev), pct_nbr_recovery_1997)) +
#   geom_boxplot() +
#   ylim(c(0.2,1.5))
```



Showing the perecent NBR recovery by burn severity class (30 years after fire)
```{r}
# df %>%
#   ggplot(aes(factor(burnSev), pct_nbr_recovery_2017)) +
#   geom_boxplot() +
#   ylim(c(0.2,1.5))
```



Showing the relationship between distance (to the edge of a medium / high severity burned area) and % NBR recovery in 2017 (26 years after fire). This is commented out because it was crashing R studio.
```{r}
# pre_fire_pixels_forest <- df %>% filter(year == 1992, landcover_code %in% 41:42) %>% pull(pixelID) 
# 
# df %>%
#   filter(pixelID %in% pre_fire_pixels_forest) %>%
#   filter(burnSev >= 3) %>%
#   ggplot(aes(distance, pct_nbr_recovery_2017)) +
#   geom_point() +
#   scale_x_continuous() +
#   scale_y_continuous() +
#   xlim(c(0,500)) +
#   adams_theme
```


plot % tree cover over time for a random sample of clusters
```{r}
#create data frame of the change in forest cover for each cluster
length(clusters_with_more_than8_pixels)
patches <- sample(clusters_with_more_than8_pixels,100)

df %>%
  filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  drop_na(land_cover_name) %>%
  group_by(clusterID,year) %>%
  summarise(pct_tree = sum(landcover_code %in% 41:43) / length(landcover_code)) %>%
  filter(clusterID %in% patches) %>%
  ggplot(aes(year,pct_tree,color = clusterID)) +
  geom_point() +
  geom_line() +
  adams_theme +
  theme(legend.position = "none") +
  scale_x_continuous(n.breaks = 7, limits = c(1992,2016)) +
  ylab(label = "Proportional Area Trees")
```


How many patches saw an increase in tree coverage during their regeneration trajectory? This measures the change in proportional area that trees take.
```{r}
df %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  #filter(WID > 0) %>%
  drop_na(land_cover_name) %>%
  group_by(clusterID,year) %>%
  summarise(pct_tree = sum(landcover_code %in% 41:43) / length(landcover_code),
            n_pixels = length(unique(pixelID))) %>%
  group_by(clusterID) %>%
  summarise(pct_change_tree = pct_tree[year==2016] - pct_tree[year==1992],
            n_pixels = head(n_pixels,1)) %>%
  ggplot(aes(pct_change_tree)) +
  geom_histogram() +
  adams_theme +
  xlab(label = "% change tree cover (2017-1992)")

#A_mean = mean(values[value_type=="A"])
```


```{r}

```



Calculate the patch-level independent variable (% change in forest cover)
```{r}
patch_level_percent_change_in_forest_cover <- df %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  drop_na(land_cover_name) %>%
  group_by(clusterID,year) %>%
  summarise(pct_tree = sum(landcover_code %in% 41:43) / length(landcover_code),
            n_pixels = length(unique(pixelID))) %>%
  group_by(clusterID) %>%
  summarise(pct_change_tree = pct_tree[year==2016] - pct_tree[year==1992],
            n_pixels = head(n_pixels,1))

names(patch_level_percent_change_in_forest_cover)

```


Calculate patch-level continuous predictor vars
```{r}
patch_level_continuous_vars <- df %>%
  select(clusterID, pixelID, burnSev,northness,eastness,elevation,slope,preFireNBR, distance) %>%
  #mutate_at(vars(ecotype_code, factsTreatmentCode, WID),as.factor) %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  group_by(clusterID) %>%
  summarise_if(is.numeric,mean)
names(patch_level_continuous_vars)
```


Calculate patch-level categorical predictor vars
```{r}
getmode <- function(x) {
   uniqv <- unique(x)
   uniqv[which.max(tabulate(match(x, uniqv)))]
}

patch_level_categorical_vars <- df %>%
  select(clusterID,pixelID,ecotype_code, eco_type, factsTreatmentCode, WID) %>%
  mutate_at(vars(ecotype_code, eco_type, factsTreatmentCode, WID),as.factor) %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  group_by(clusterID) %>%
  summarise_if(is.factor,getmode) 

names(patch_level_categorical_vars)
```


Joining patch-level dataframes
```{r}
patch_data  <- patch_level_percent_change_in_forest_cover %>%
  left_join(patch_level_continuous_vars, by = "clusterID") %>%
  left_join(patch_level_categorical_vars, by = "clusterID") %>%
  select(-clusterID, -n_pixels)

patch_data
```



Exploring correlations between variables
```{r}
library(ggcorrplot)

corr <- patch_data %>%
  select_if(is.numeric) %>%
  cor() %>% round(1)

Pmat <- patch_data %>%
  select_if(is.numeric) %>%
  cor_pmat()

ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, p.mat = Pmat)
```



```{r}
patch_data %>%
  ggplot(aes(x = eco_type,y = pct_change_tree)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```







Extra
```{r}

#representation of starting (pre-fire) physiognomic classes
df %>%
  filter(year == 1992) %>%
  group_by(land_cover_name) %>%
  summarise(number_pixels = length(unique(pixelID))) %>%
  drop_na() %>%
  mutate(sample_area_hectares = number_pixels * 900 / 10000) %>%
  ggplot(aes(land_cover_name, number_pixels)) +
  geom_bar(stat = "identity") +
  adams_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


#representation of ending (post-fire) physiognomic classes
df %>%
  filter(year == 2016) %>%
  group_by(land_cover_name) %>%
  summarise(number_pixels = length(unique(pixelID))) %>%
  drop_na() %>%
  mutate(sample_area_hectares = number_pixels * 900 / 10000) %>%
  ggplot(aes(land_cover_name, number_pixels)) +
  geom_bar(stat = "identity") +
  adams_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


Analysis of the effect of planting trees on the future physiognomic class in fire affected pixels.


First look at % transitioning in treated and untreated pixels
```{r}
nonforest1992 <- df %>% 
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  filter(year == 1992, landcover_code %notin% 41:43) %>% pull(pixelID)
forest2016 <- df %>% 
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  filter(year == 2016, landcover_code %in% 41:43) %>% pull(pixelID)
transition2forest <- unique(df$pixelID[(df$pixelID %in% nonforest1992) & (df$pixelID %in% forest2016)])

transition2forest

df %>%
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  mutate(transitioned = pixelID %in% transition2forest) %>%
  mutate(planting = factsTreatmentCode %in% c(4431,4432)) %>%
  group_by(pixelID) %>%
  summarise(transitioned = max(transitioned), planted = max(planting)) %>%
  group_by(planted) %>%
  summarise(percent_trans = sum(transitioned) / length(transitioned), n = length(transitioned)) %>%
  ggplot(aes(planted,percent_trans)) +
  geom_bar(stat = "identity")
```


Pixel-level data, difference of differences (i.e. t-test)
```{r}
clusters_with_treatment <- df %>%
  select(clusterID, pixelID, year, landcover_code, factsTreatmentCode) %>%
  mutate(transitioned = pixelID %in% transition2forest) %>%
  mutate(planting = factsTreatmentCode %in% c(4431,4432)) %>%
  group_by(pixelID) %>% summarise(clusterID = first(clusterID), tr =   first(transitioned), pl = max(planting)) %>%
  group_by(clusterID,pl) %>% summarise(n_tr = sum(tr), n = length(tr)) %>%
  group_by(clusterID) %>%
  summarise(nb = length(n)) %>%
  filter(nb > 1) %>% pull(clusterID)

facts_treatment_df <- df %>%
  select(clusterID, pixelID, year, landcover_code, factsTreatmentCode) %>%
  mutate(transitioned = pixelID %in% transition2forest) %>%
  mutate(planting = factsTreatmentCode %in% c(4431,4432)) %>%
  #filter(clusterID %in% cluster_with_treatment) %>%
  group_by(pixelID) %>% summarise(clusterID = first(clusterID), tr = first(transitioned), pl = max(planting)) %>%
  group_by(clusterID,pl) %>% summarise(n_tr = sum(tr), n = length(tr)) %>%
  mutate(frac_tr = n_tr / n) %>%
  mutate(lnfrac_tr = log(frac_tr + 0.000001))

#treatGroup <- plantedVSnot %>% filter(pl == 1) %>% pull(frac_tr)

#ControlGroup <- plantedVSnot %>% filter(pl == 0) %>% pull(frac_tr)

```

Visualizing data 
```{r}

facts_treatment_df %>%
  ggplot(aes(factor(pl),frac_tr)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=3) 

```

```{r}

facts_treatment_df %>%
ggplot(aes(frac_tr)) +
geom_histogram()

facts_treatment_df %>%
ggplot(aes(log(frac_tr))) +
geom_histogram()
```


```{r}
clusters_with_facts_treatment <- facts_treatment_df %>%
  filter(clusterID %in% clusters_with_treatment)  %>% 
  group_by(clusterID) %>%
  summarise(nb = length(n)) %>%
  filter(nb > 1) %>% pull(clusterID)

facts_treatment_df <- facts_treatment_df %>%
  filter(clusterID %in% clusters_with_facts_treatment)


t.test(lnfrac_tr ~ pl, data = test, paired = TRUE)
```



