---
title: "analysis_script.Rmd"
author: "DS421 Team"
date: "March 9, 2020"
output: md_document
---
Install Packages
```{r}
library(tidyverse)
source('scripts/adams_theme.R') #for ggplot2
```


Import data
```{r}
df <- read_csv('data/FireDataKlamath1987part1_v2.csv', col_types = cols(clusterID = col_factor())) %>%
  rbind(read_csv('data/FireDataKlamath1987part2_v2.csv', col_types = cols(clusterID = col_factor()))) %>%
  rbind(read_csv('data/FireDataKlamath1987part3_v2.csv', col_types = cols(clusterID = col_factor())))


#this csv describe each ecotype code
eco_type_codes <- read_csv('data/landfire_ecotype_codes.csv')

n_pixels <- nrow(df) #number of pixels in the dataset 
n_pixels
```

Some initial cleaning of the data
```{r}
df <- df %>%
  select(-.geo, -'system:index') %>%
  mutate(pixelID = as.character(1:n_pixels))

names(df)
```


Identifying patches with more than eight pixels. "Clusters" and "Patches" are used interchangeably throughout the document.
```{r}
clusters_with_more_than8_pixels <- df %>% group_by(clusterID) %>%
  summarise(Npixels = length(unique(pixelID))) %>%
  filter(Npixels > 8) %>%
  pull(clusterID)
```


Creating a data frame that has the landcover classifications of each pixel in years where the NLCD landcover data is available.
```{r}

landcover <- names(df)[grep(names(df), pattern = "NLCD")]

df_lc <- df %>%
  select(pixelID, clusterID, landcover) %>%
  gather(landcover, key = "NLCD_year", value = "landcover_code") %>%
  mutate(land_cover_name = case_when(
    (landcover_code == 71) ~ "grass_herb",
    (landcover_code == 51) ~ "shrub",
    (landcover_code == 42) ~ "evergreen_forest",
    (landcover_code == 43) ~ "mixed forest",
    (landcover_code == 31) ~ "barren_land",
    (landcover_code == 23) ~ "developed",
    (landcover_code == 41) ~ "deciduous_forest",
    (landcover_code == 11) ~ "open_water",
    (landcover_code == 81) ~ "pasture_hay",
    (landcover_code == 33) ~ "transitional_barren",
    (landcover_code == 52) ~ "shrub",
    (landcover_code == 21) ~ "developed",
    (landcover_code == 95) ~ "emergent_herb_wetlands",
    (landcover_code == 90) ~ "woody_wetlands",
    (landcover_code == 22) ~ "developed"
  )) %>%
  mutate(year = as.numeric(gsub("NLCD", "", NLCD_year))) %>%
  arrange(clusterID)

df_lc
```


Creating a tree canopy cover df in years where % tree canopy cover data is available
```{r}

TCC <- names(df)[grep(names(df), pattern = "TCC")]

df_tcc <- df %>%
  select(pixelID, clusterID, TCC) %>%
  gather(TCC, key = "TCCyear", value = "pct_TCC") %>% 
  mutate(year = as.numeric(gsub("TCC", "", TCCyear)))

df_tcc
```


Creating an NBR dataframe showing the NBR value in each pixel for each year.
```{r}
years <- as.character(1985:2017)
df_nbr <- df %>%
  select(pixelID, clusterID, years) %>%
  gather(years, key = "year_string", value = "NBR") %>%
  mutate(year = as.numeric(year_string)) %>%
  select(-year_string, -clusterID) 

df_nbr 
```


Creating a FACTS df showing which pixels were planted with trees after the 1987 fires
```{r}

facts_cols <- names(df)[grep(names(df), pattern = "facts")]
df_facts <- df %>%
  select(pixelID, clusterID, facts_cols) %>%
  gather(facts_cols, key = "facts_year", value = "factsTreatmentCode") %>%
  mutate(factsTreatment = case_when(
    (factsTreatmentCode == 0) ~ "none",
    (factsTreatmentCode == 4431) ~ "plant_trees",
    (factsTreatmentCode == 4432) ~ "replant_or_fill"
  )) %>%
  mutate(year = as.numeric(gsub("facts", "", facts_year))) %>%
  arrange(clusterID)

df_facts
```


Merging dfs together
```{r}
#mering dfs together
df <- df %>% select(pixelID, clusterID, FIRE_YEAR, FireID, burnSev, distance, northness, eastness, elevation, slope, b1, WID) %>%
  rename(ecotype_code = b1) %>%
  left_join(df_nbr, by = c("pixelID")) %>%
  left_join(df_lc,c("pixelID","year")) %>%
  left_join(df_facts,c("pixelID","year")) %>% 
  mutate(burnSevLong = case_when(
    (burnSev == 0) ~ "background/no_data",
    (burnSev == 1) ~ "verylow/no_burn",
    (burnSev == 2) ~ "low",
    (burnSev == 3) ~ "med",
    (burnSev == 4) ~ "high"
  )) %>% 
  mutate(eco_type = case_when(
    (ecotype_code == 549) ~ "MC_LM_Serp",
    (ecotype_code == 550) ~ "MC_UM_Serp",
    (ecotype_code == 551) ~ "MC_DrytoMes",
    (ecotype_code == 552) ~ "MC_Mes",
    (ecotype_code == 553) ~ "MixedOak",
    (ecotype_code == 554) ~ "MC_BO_LM",
    (ecotype_code == 555) ~ "Jeffrey_Ponderosa",
    (ecotype_code == 556) ~ "RedFir",
    (ecotype_code == 557) ~ "SubalpineWoodland",
    (ecotype_code == 558) ~ "Maritime_DFir_WHem"
  )) %>%
  filter(ecotype_code %in% eco_type_codes$eco_type_code) 

df
#adding the ecotype descriptions
```



Checking representation of forest types in our sample of pixels. The majority are "mixed conifer dry to mesic", and "mixed conifer mesic".
```{r}
#representation of forest types
df %>%
  group_by(eco_type) %>%
  summarise(number_pixels = length(unique(pixelID))) %>%
  drop_na() %>%
  mutate(sample_area_hectares = number_pixels * 900 / 10000) %>%
  ggplot(aes(eco_type, number_pixels)) +
  geom_bar(stat = "identity") +
  adams_theme +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab(label = "N pixels")



total_sample_area <- length(unique(df$pixelID)) * 900 / 10e4 
total_sample_area 

total_number_patches <- length(unique(df$clusterID))
total_number_patches
```


Exploring distribution of patch size in units of hectares and pixels in our sample
```{r}
df %>%
  group_by(clusterID) %>%
  summarise(num_pixels_per_cluster = length(unique(pixelID))) %>%
  mutate(cluster_area = num_pixels_per_cluster * 900 / 10000) %>%
  pull(cluster_area) %>% summary()
  
df %>%
  group_by(clusterID) %>%
  summarise(num_pixels_per_cluster = length(unique(pixelID))) %>%
  mutate(cluster_area = num_pixels_per_cluster * 900 / 10000) %>%
  ggplot(aes(num_pixels_per_cluster)) +
  geom_histogram() +
  xlim(c(0,100)) +
  adams_theme

length(unique(df$pixelID))
length(unique(df$pixelID)) * 900 / 10e4
  
```


Quantifying how many pixels went from non-forest back to forest in their post-fire regeneration trajectory. These are the pixels of interest for this analysis because we are not interested in pixels that were never denuded of forest.
```{r}
`%notin%` <- Negate(`%in%`)

nonForest2001 <- df %>%
  filter(year == 2001, landcover_code %notin% 41:43) %>%
  pull(pixelID) %>% unique()

#how many pixels went from shrub to tree
df %>%
  filter(pixelID %in% nonForest2001) %>%
  filter(year == 2016, landcover_code %in% 41:43) %>%
  pull(pixelID) %>% unique() %>% length()

```

Plotting the fraction of pixels in each patch that were classified as forest in multiple years since the time of fire. The below plot shows these fractions from 1992 to 2016 (the fire was in 1987).
```{r}
#create a data frame of the change in forest cover for each patch (cluster = patch)
length(clusters_with_more_than8_pixels) # number of patches in the sample. Patches with less than 8 pixels were excluded.
patches <- sample(clusters_with_more_than8_pixels,100)

df %>%
  filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  drop_na(land_cover_name) %>%
  group_by(clusterID,year) %>%
  summarise(pct_tree = sum(landcover_code %in% 41:43) / length(landcover_code)) %>%
  filter(clusterID %in% patches) %>%
  ggplot(aes(year,pct_tree, group = clusterID)) +
  geom_point() +
  geom_line() +
  adams_theme +
  theme(legend.position = "none") +
  scale_x_continuous(n.breaks = 7, limits = c(1992,2016)) +
  ylab(label = "Tree Cover Fraction Per Patch")
```


Whats the distribution of changes in the fractional area of tree cover between 2016 and 1992 for the patches in our sample? Below we see that most patches did not change their area of tree cover, while some accumulated tree cover and some lost tree cover.


```{r}
#Calculating the patch-level dependent variable (% change in tree cover)
patch_level_percent_change_in_forest_cover <- df %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  drop_na(land_cover_name) %>%
  group_by(clusterID,year) %>%
  summarise(pct_tree = sum(landcover_code %in% 41:43) / length(landcover_code),
            n_pixels = length(unique(pixelID))) %>%
  group_by(clusterID) %>%
  summarise(pct_change_tree = pct_tree[year==2016] - pct_tree[year==1992],
            n_pixels = head(n_pixels,1))

#plotting the distribution of changes in the fractional area of tree cover between 2016 and 1992 for the patches in our sample
patch_level_percent_change_in_forest_cover %>%
 ggplot(aes(pct_change_tree)) +
  geom_histogram() +
  adams_theme +
  xlab(label = "% change in tree cover (1992-2016)")
```

Pre-fire NBR is a measure of how much vegetation was in the pixel before the fire.
```{r}
#creating pre-fire NBR values for each pixel
pre_fire_nbr <- df %>% 
  filter(year %in% 1985:1986) %>%
  group_by(pixelID) %>%
  summarise(preFireNBR = mean(NBR))

#adding the percent NBR recovery to the df
df <- df %>%
  left_join(pre_fire_nbr) %>%
  mutate(pct_nbr_recovery = NBR / preFireNBR)
```


Calculating the patch-level continuous predictor vars
```{r}
patch_level_continuous_vars <- df %>%
  select(clusterID, pixelID, burnSev,northness,eastness,elevation,slope,preFireNBR, distance) %>%
  #mutate_at(vars(ecotype_code, factsTreatmentCode, WID),as.factor) %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  group_by(clusterID) %>%
  summarise_if(is.numeric,mean)
names(patch_level_continuous_vars)
```


Calculating the patch-level categorical predictor vars
```{r}
getmode <- function(x) {
   uniqv <- unique(x)
   uniqv[which.max(tabulate(match(x, uniqv)))]
}

patch_level_categorical_vars <- df %>%
  select(clusterID,pixelID,ecotype_code, eco_type, factsTreatmentCode, WID) %>%
  mutate_at(vars(ecotype_code, eco_type, factsTreatmentCode, WID),as.factor) %>%
  #filter(burnSev > 1) %>%
  filter(clusterID %in% clusters_with_more_than8_pixels) %>%
  group_by(clusterID) %>%
  summarise_if(is.factor,getmode) 

#instead of calculating the mode of the facts treatments in each patch, here we create another variable that says a patch is "treated" (i.e. logical value of `1`) if any pixel in the patch was treated.

facts_treatment_df_v2 <- df %>%
  select(clusterID,pixelID,factsTreatmentCode) %>%
  group_by(clusterID) %>%
  summarise(factsTreatment = max(factsTreatmentCode)) %>%
  mutate(factsTreatmentLogical = factor(as.numeric(factsTreatment > 1)))

patch_level_categorical_vars <-patch_level_categorical_vars %>%
  left_join(facts_treatment_df_v2)

#names(patch_level_categorical_vars)
```


Joining patch-level dataframes to have the continuous predictor vars, the categorical predictor vars, and the dependent variables in one df.
```{r}
patch_data  <- patch_level_percent_change_in_forest_cover %>%
  left_join(patch_level_continuous_vars, by = "clusterID") %>%
  left_join(patch_level_categorical_vars, by = "clusterID") %>%
  select(-clusterID, -n_pixels)

patch_data
```







```{r}
names(df)
```





Exploring correlations between variables
```{r}
library(ggcorrplot)

corr <- patch_data %>%
  select_if(is.numeric) %>%
  cor() %>% round(1)

Pmat <- patch_data %>%
  select_if(is.numeric) %>%
  cor_pmat()

ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, p.mat = Pmat)
```


Analyzing forest recovery by forest type. The majority of forest types continue to lose forest cover even 5 years after the fire. The Mixed Conifer Mesic Forest type appears to have the most promising forest regeneration.
```{r}
patch_data %>%
  ggplot(aes(x = eco_type,y = pct_change_tree)) +
  geom_boxplot() +
  ylab(label = "% change in tree cover (1992-2016)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
```


Analyzing the effect of the USFS planting trees on the future physiognomic class in fire-affected pixels using data from the Forest Activity Tracking System dataset.

A first look at % of pixels transitioning from non-forest to forest in treated (planted) and untreated pixels.
```{r}
nonforest1992 <- df %>% 
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  filter(year == 1992, landcover_code %notin% 41:43) %>% pull(pixelID)
forest2016 <- df %>% 
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  filter(year == 2016, landcover_code %in% 41:43) %>% pull(pixelID)
transition2forest <- unique(df$pixelID[(df$pixelID %in% nonforest1992) & (df$pixelID %in% forest2016)])

#transition2forest
df %>%
  select(pixelID, year, landcover_code, factsTreatmentCode) %>%
  mutate(transitioned = pixelID %in% transition2forest) %>%
  mutate(planting = factsTreatmentCode %in% c(4431,4432)) %>%
  group_by(pixelID) %>%
  summarise(transitioned = max(transitioned), planted = max(planting)) %>%
  group_by(planted) %>%
  summarise(percent_trans = sum(transitioned) / length(transitioned), n = length(transitioned)) %>%
  ggplot(aes(factor(planted),percent_trans)) +
  geom_bar(stat = "identity") +
  ylab(label = "percent of pixels regenerating back to forest") +
  xlab(label = "not planted = 0, planted = 1")
```


Setting up a datafram to test the difference in differences between paired control and planted pixels. Groups of pixels that are in the same patch, but that differ in their treatement (planted vs. not planted) are matched.
```{r}
facts_treatment_df <- df %>%
  filter(year > 1986) %>%
  select(clusterID, pixelID, year, landcover_code, factsTreatmentCode) %>%
  mutate(transitioned = pixelID %in% transition2forest) %>%
  mutate(planting = factsTreatmentCode %in% c(4431,4432)) %>%
  #filter(clusterID %in% cluster_with_treatment) %>%
  group_by(pixelID) %>% summarise(clusterID = first(clusterID), tr = first(transitioned), pl = max(planting)) %>%
  group_by(clusterID,pl) %>% summarise(n_tr = sum(tr), n = length(tr)) %>%
  mutate(frac_tr = n_tr / n) %>%
  mutate(lnfrac_tr = log(frac_tr + 0.000001)) %>%
  mutate(treatment = case_when(
    (pl == 0 ~ "control"),
    (pl == 1 ~ "planted")
  ))

#treatGroup <- plantedVSnot %>% filter(pl == 1) %>% pull(frac_tr)

#ControlGroup <- plantedVSnot %>% filter(pl == 0) %>% pull(frac_tr)

```

Visualizing data 
```{r}
facts_treatment_df %>%
  ggplot(aes(factor(treatment),frac_tr)) +
  geom_boxplot() +
  stat_summary(fun.y = mean, geom="point",colour="darkred", size=3) +
  ylab("percent of pixels regenerating back to forest") +
  xlab("treatment group")+
  labs(title = "the effect of planting on patch-level forest regeneration (30 yrs post fire)")

```

The dependent variable (y axis in figure above) needs to be log transformed for statistical analysis because it is not normally distributed. A log transform resolves this.
```{r}
facts_treatment_df %>%
ggplot(aes(frac_tr)) +
geom_histogram()

facts_treatment_df %>%
ggplot(aes(log(frac_tr))) +
geom_histogram()
```

A paired t-test shows that planting does increase forest regeneration 30 years after fire.
```{r}
clusters_with_treatment <- facts_treatment_df %>%
  drop_na(clusterID) %>%
  group_by(clusterID) %>%
  summarise(nb = length(n)) %>%
  filter(nb > 1) %>% pull(clusterID)

facts_treatment_df %>% filter(clusterID %in% clusters_with_treatment)

t.test(lnfrac_tr ~ pl, 
       data = facts_treatment_df %>% filter(clusterID %in% clusters_with_treatment),
       paired = TRUE)
```


But, in what scenarios does planting help the most compared to control? The correlation plot below suggests that "tr_diff" (the difference in the % of pixels transitioning back to forest between the treatment group and the paired control group) is positively correlated with slope, eastness, and pre-Fire NBR value. Therefore, we can assume that planting helps regeneration the most on steeper slopes, east-facing slopes, and in locations where there was thicker vegetation before the fire. This last variable could be a proxy for "site potential".
```{r}
facts_cor_data <- facts_treatment_df %>% 
  filter(clusterID %in% clusters_with_treatment) %>%
  select(-n_tr, -n,-lnfrac_tr, -pl) %>%
  spread(key = treatment, value = frac_tr) %>%
  mutate(tr_diff = planted - control) %>%
  select(-control,-planted) %>%
  left_join(patch_level_continuous_vars, by = "clusterID") %>%
  left_join(patch_level_categorical_vars, by = "clusterID") %>%
  ungroup %>%
  select(-factsTreatmentCode, -clusterID) %>%
  select_if(is.numeric) %>%
  drop_na()

corr <- facts_cor_data %>%
  select_if(is.numeric) %>%
  cor() %>% round(1)

Pmat <- facts_cor_data %>%
  select_if(is.numeric) %>%
  cor_pmat()

ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, p.mat = Pmat)
```



```{r}
facts_cor_data %>%
  ggplot(aes(x = preFireNBR, y = tr_diff)) +
  geom_point()
```






##Appendix. Exploring NBR recovery over time.
Adding the % NBR recovey to the df. %NBR recovery is the fraction of mean pre-fire NBR

Showing the distribution of NBR values for pixels classified as forest for NLCD
```{r}

df %>%
  drop_na(land_cover_name) %>%
  filter(landcover_code %in% c(41:42, 71,51,52)) %>%
  ggplot(aes(x=NBR, fill=land_cover_name)) +
  geom_density( color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_manual(values=c("light green", "dark green", "yellow", "brown")) +
  #theme_ipsum() +
  labs(fill="") +
  adams_theme


```


Plotting the patch-level NBR through time for a random sample of patches. We see that NBR starts to saturate (i.e. reach the pre-fire NBR value) within about 10-15 years. This is despite the fact that pixels have not actually returned to forest. This is evidence that using NBR is not suitable of measuring forest recovery.
```{r}
#patches <- sample(unique(df$clusterID),100)

df %>%
  group_by(FireID, clusterID, year) %>%
  summarise_if(is.numeric,mean) %>%
  #filter(clusterID %in% patches) %>%
  ggplot(aes(year,NBR,color = clusterID)) +
  geom_line() +
  adams_theme +
  theme(legend.position = "none")
```


Plotting the percent NBR recovery for each patch. This shows NBR as a percent of the pre-fire values.
```{r}
#patches <- sample(unique(df$clusterID),10)
df %>%
  group_by(clusterID, year) %>%
  #filter(WID > 0) %>%
  summarise_if(is.numeric,mean) %>%
  #filter(clusterID %in% patches) %>%
  ggplot(aes(year,pct_nbr_recovery,color = clusterID)) +
  geom_line() +
  ylim(c(-1,1.5)) +
  adams_theme +
  theme(legend.position = "none")
```



Histogram of percent recovery in 2000 (6 years after fire)
```{r}
df %>%
  group_by(pixelID, year) %>%
  summarise_if(is.numeric,mean) %>%
  filter(year == 1992) %>%
  ggplot(aes(x = pct_nbr_recovery, fill = factor(burnSev))) +
  geom_density(alpha=0.6, position = 'identity') +
  scale_fill_brewer(palette = "Reds") +
  xlim(c(-0.2,2)) +
  xlab(label = "% NBR Recovery 5 years post-fire") +
  adams_theme +
  labs(fill="Burn Severity") 
  
```

Density plot of percent recovery in 2017 (23 years after fire)
```{r}
df %>%
  group_by(pixelID, year) %>%
  summarise_if(is.numeric,mean) %>%
  filter(year == 2017) %>%
  ggplot(aes(x = pct_nbr_recovery, fill = factor(burnSev))) +
  geom_density(alpha=0.6, position = 'identity') +
  scale_fill_brewer(palette = "Reds") +
  xlim(c(-0.2,2)) +
  xlab(label = "% NBR Recovery 30 years post-fire") +
  adams_theme +
  labs(fill="Burn Severity") 
```


Creating variables showing the percent NBR recovery in 92,97,2017
```{r}
df <- df %>%
  filter(pct_nbr_recovery >= -10)

pctNBRrec2017 <- df %>%
  filter(year == 2017) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_2017 = pct_nbr_recovery)

pctNBRrec1997 <- df %>%
  filter(year == 1997) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_1997 = pct_nbr_recovery)

pctNBRrec1992 <- df %>%
  filter(year == 1992) %>%
  select(pixelID, pct_nbr_recovery) %>%
  rename(pct_nbr_recovery_1992 = pct_nbr_recovery)

df <- df %>%
  left_join(pctNBRrec2017) %>%
  left_join(pctNBRrec1997) %>%
  left_join(pctNBRrec1992)
```



Showing the perecent NBR recovery by burn severity class (10 years after fire)
```{r}
# df %>%
#   ggplot(aes(factor(burnSev), pct_nbr_recovery_1997)) +
#   geom_boxplot() +
#   ylim(c(0.2,1.5))
```



Showing the perecent NBR recovery by burn severity class (30 years after fire)
```{r}
# df %>%
#   ggplot(aes(factor(burnSev), pct_nbr_recovery_2017)) +
#   geom_boxplot() +
#   ylim(c(0.2,1.5))
```



Showing the relationship between distance (to the edge of a medium / high severity burned area) and % NBR recovery in 2017 (26 years after fire). This is commented out because it was crashing R studio.
```{r}
# pre_fire_pixels_forest <- df %>% filter(year == 1992, landcover_code %in% 41:42) %>% pull(pixelID) 
# 
# df %>%
#   filter(pixelID %in% pre_fire_pixels_forest) %>%
#   filter(burnSev >= 3) %>%
#   ggplot(aes(distance, pct_nbr_recovery_2017)) +
#   geom_point() +
#   scale_x_continuous() +
#   scale_y_continuous() +
#   xlim(c(0,500)) +
#   adams_theme
```


