---
title: "fire_analysis"
author: "Adam Hanbury-Brown"
date: "April 21, 2020"
output: html_document
---

---
title: "Hypothesis 2"
author: "Adam Hanbury-Brown"
date: "April 21, 2020"
output: html_document
---

##Hypothesis 2
This script tests hypothesis two: A combination of topography, climate, fire, and management variables will predict the probability of conifer canopy dominance 30 yrs after fire. 

The dependent variable is boolean: whether or not a post-fire pixel is classified as being confier or not conifer (using a random forest algorithm).

**Fire year**: 1987
**Study area**: The Klamath Ecoregion
**Pixels included**: Mixed conifer forest as classified by Landfire's biophysical settings, identified as being conifer before the fire (as classified from a random forest algorithm), that burned at medium or high severity (MTBS), and that did not subsequently burn. Essentially, the study pixels included experienced stand-replacing fires in 1987 that occured in mixed conifer forest.






Install Packages
```{r}
library(tidyverse)
library(foreign)
source('scripts/adams_theme.R')
source('scripts/adams_theme_v2.R')#for ggplot2
library(sf)
library(caret)
```

Import data
```{r}
df <- readRDS('data/cleanDataApril20_v3.Rds')
#this csv describe each ecotype code

str(df)
eco_type_codes <- read_csv('data/landfire_ecotype_codes.csv')
landfire_codes <- foreign::read.dbf("data/US_200EVT\\US_200EVT.dbf")
landfire_codes <- landfire_codes %>% dplyr::select(VALUE,EVT_PHYS) %>% rename(landfire2016 = VALUE)
n_pixels <- nrow(df) #number of pixels in the dataset 
n_pixels
```


```{r}
#cleaning
df <- df %>%
  left_join(landfire_codes) %>%
  mutate(conifer = case_when(
    (EVT_PHYS == "Conifer" | EVT_PHYS == "Conifer-Hardwood") ~ TRUE,
    TRUE ~ FALSE
  ))
#str(df)
```

exporting fields to make field descriptions
```{r}
#names(df) %>% as.tibble() %>% write_csv(path = "data/data_fields.csv")
```

creating dataframe at the SUID level
```{r}
# SUID_df <- df %>%
#   mutate_at(.vars = "SUID", .funs = as.factor) %>%
#   drop_na(SUID) %>%
#   group_by(SUID) %>%
#   summarise_if(.predicate = is.numeric, .funs = mean) 
```


defining and reclassing variables 
```{r}
continuous_predvars <- names(df)[c(6:16)]
categorical_predvars <- names(df)[c(4,5,27,28)]
boolean_predvars <- names(df)[c(19,21,22,23,24,26)]
depVars <- names(df)[c(25,29,30)]

df <- df %>%
  mutate_at(.vars = vars(boolean_predvars), .funs = as.logical) %>%
  mutate_at(.vars = vars(categorical_predvars), .funs = as.factor) %>%
  mutate_at(.vars = "SUID",.funs = as.factor)

str(df)
```


51,521 pixels were not in a SUID. 14,241 pixels were in a SUID (21 % of the study area). Total study area is about 6,000 hectares, with about 1,200 ha of managed land in the study. There are 360 SUIDs in the study area.
```{r}
#percent of study area in a SUID
(nrow(df) - sum(is.na(df$SUID))) / nrow(df) * 100

#total study area (ha)
nrow(df) * 900 / 10000

#number of SUIDs in the study area
length(unique(df$SUID))
```



Summary of continuous variables
```{r}
for(i in continuous_predvars){
  print(i)
  print(list((df %>% pull(i) %>% summary())))
}
```


Export raster of landfire codes
```{r}

lf_lookup_code <- landfire_codes %>% pull(EVT_PHYS) %>% unique() %>% as.tibble() %>% rowid_to_column(var = "lf_code")

lf_code <- landfire_codes %>% pull(EVT_PHYS) %>% unique() %>% as.tibble() %>% rowid_to_column(var = "lf_code") %>% rename(EVT_PHYS = value) %>%
  left_join(df) %>% dplyr::select(x,y,lf_code)

summary(lf_code$lf_code)

library(raster)

SUID_polygons <- st_read(dsn = "~/cloud/gdrive/fire_project/local_data/facts/SUIDSin1987StudyArea.shp")


lf_coderaster <- rasterFromXYZ(xyz = lf_code, res = 30, crs = crs(SUID_polygons))
plot(lf_coderaster)

#writeRaster(lf_coderaster,filename = "~/cloud/gdrive/fire_project/local_data/Landfire_EVC/lf.tiff")



```



Looking at collinearity among the continuous variables
```{r}
library(ggcorrplot)
library(tidyverse)

corr <- df %>%
  dplyr::select(continuous_predvars) %>%
  cor() %>% round(2)

Pmat <- df %>%
  dplyr::select(continuous_predvars) %>%
  cor_pmat()

ggcorrplot(corr, hc.order = TRUE, type = "lower",
   lab = TRUE, p.mat = Pmat)
```


Correlation among categorical variables
```{r}
categorical_predvars
table(df$conifer,df$burnSev)
chisq.test(df$burnSev,df$forestType)

```



Exploring relationships visually

The SAP surviving conifers appears to be higher for the pixels that were conifer in 2016. The same is not true of the SAP_burn
```{r}
df %>%
  ggplot(aes(x = conifer, y = SAP_surv_conifers))+
  geom_boxplot()
```



More pixels in medium severity burn sites become conifer in 2016
```{r}

print('burn sev')
table(df$conifer,df$burnSev)
chisq.test(df$conifer,df$burnSev)

print('planting')
table(df$conifer,df$plantedFirst5Years)
chisq.test(df$conifer,df$plantedFirst5Years)


```


Solar load is higher for the pixels that did not come back as conifer
```{r}
source('~/cloud/gdrive/fire_project/DS421_proj/scripts/adams_theme_v2.R')#for ggplot2

northness_plot <- df %>%
  ggplot(aes(x = conifer, y = northness))+
  geom_boxplot() + 
  ylab(label = "Northness (cos of aspect)") +
  xlab(label = "Conifers dominate canopy 30 yrs later") +
  adams_theme2 

  
png(filename = "~/cloud/gdrive/fire_project/DS421_proj/output/northness.png", height=5, width=8, units="in", res = 100)
northness_plot
dev.off()



SAP_figure <- df %>%
  ggplot(aes(x = conifer, y = SAP_surv_conifers))+
  geom_boxplot() + 
  ylab(label = "Seed availability proxy") +
  xlab(label = "Conifers dominate canopy 30 yrs later") +
  adams_theme2 

png(filename = "~/cloud/gdrive/fire_project/DS421_proj/output/SAP_figure.png", height=5, width=8, units="in", res = 100)
SAP_figure 
dev.off()
  
```




Northness is higher for the pixels that did not come back as conifer. Eastness doesn't matter as much.
```{r}
df %>%
  ggplot(aes(x = conifer, y = northness))+
  geom_boxplot()

df %>%
  ggplot(aes(x = conifer, y = SAP_surv_conifers))+
  geom_boxplot()

table(df$conifer,df$SPB, dnn = c("con","SPB"))

```



```{r}
df %>%
  ggplot(aes(x = CWD))+
  geom_histogram()

df %>%
  ggplot(aes(x = conifer, y = CWD))+
  geom_boxplot()

df %>%
  ggplot(aes(x = conifer, y = elevation))+
  geom_boxplot()

df %>%
  ggplot(aes(x = conifer, y = slope))+
  geom_boxplot()
```


Testing VIF on a model I hypothesize to explain the probability that a pixel becomes conifer. VIF is not over 5 for any variable.
```{r}
library(tidyverse)
library(car)


df <- df %>% mutate_at(.vars = "AvGTemp",.funs = function(x){x/12}) %>%
  filter(forestType != "550")

#Create training set
train <- df %>% sample_frac(.70)

table(train$conifer)


#Create test set
test  <- anti_join(df, train, by = 'ID')
test


#grab equal rows of conifer = T and conifer = F
train1 <- train %>% filter(conifer == F)
train2 <- train %>% filter(conifer == T) %>% sample_n(size = 3449)
train <- rbind(train1,train2)

#model with all variables
attach(train)
fullMod <- glm(formula = conifer ~ SAP_surv_conifers + SAP_burn + elevation + slope + northness + eastness + CWD + AvGppt + AvGTemp + SolarLoad + burnSev + forestType + LAND_SUITA + plantedFirst5Years + plantedAfter5Yrs + SPB + SPO + SPMec + certified_plantedIn6Years, family = binomial())

#alias(fullMod)

summary(fullMod)

car::vif(fullMod)
```



Select key variables 
```{r}

Mod_rownames <- rownames(summary(fullMod)$coefficients)

print(paste0("AIC",fullMod$aic))

key_vars <- summary(fullMod)$coefficients %>% as_tibble() %>% mutate(rnames = Mod_rownames) %>% filter(`Pr(>|z|)` < 0.01)


#take out solar load because its correlated with northness and northness appears to be better

key_vars <- filter(key_vars, rnames != "SolarLoad")

key_vars

```

AIC with just these "important vars" has a way higher AIC. 
```{r}

mod2 <- glm(data = train, formula = conifer ~ elevation + northness + eastness + AvGppt + AvGTemp + burnSev + plantedFirst5Years + SPO + SPMec, family = binomial())

print(paste("AIC",mod2$aic))

mod3 <- glm(data = train,formula = conifer ~ SAP_surv_conifers + SAP_burn + elevation + slope + SolarLoad + eastness + CWD + AvGppt + AvGTemp + burnSev + forestType + LAND_SUITA + plantedFirst5Years + plantedAfter5Yrs + SPB + SPO + SPMec + certified_plantedIn6Years, family = binomial())

mod4 <- glm(data = train,formula = conifer ~ elevation + slope + northness + eastness + CWD + AvGppt + forestType + AvGTemp + burnSev + plantedFirst5Years + SPB + SPO + SPMec + certified_plantedIn6Years, family = binomial())


```

```{r}
mod5 <- glm(data = train,formula = conifer ~ elevation + northness + eastness + AvGppt + AvGTemp + burnSev + plantedFirst5Years + SPB + SPMec, family = binomial())


vif(mod4)

print(paste("AIC mod 3",mod3$aic))

print(paste("AIC mod small",mod2$aic))

print(paste("AIC full mod",fullMod$aic))


print(paste("AIC 4",mod4$aic))
print(paste("AIC 5",mod5$aic))

```



```{r}
library(sjPlot)
library(sjmisc)
library(sjlabelled)
```



making a table to bring into report and presentation
```{r}
library(xtable)

#print(xtable(summary(mod4)),type = "html")

tab_model(mod5)
```



```{r}
xtable(summary(mod5))
```


```{r}

#levels(test$forestType) <- c("549","549","551","552","555","556","559")
#levels(train$forestType) <- c("549","549","551","552","555","556","559")


outOfSampleTest <- test %>% mutate(prediction = predict(mod4, newdata = test, type = "response")) %>%
  mutate(classification = as.logical(round(prediction))) %>% dplyr::select(prediction, conifer, classification) 

#confusion matrix
table(outOfSampleTest$conifer, outOfSampleTest$classification)


#
confusionMatrix(as.factor(outOfSampleTest$conifer), as.factor(outOfSampleTest$classification))


prediction_fig <- outOfSampleTest %>%
  ggplot(aes(conifer,prediction)) +
  geom_boxplot()+ 
  ylab(label = "Predicted Prob. Conifer") +
  xlab(label = "Reference (Not Conifer vs. Conifer)") +
  adams_theme2 

  
png(filename = "~/cloud/gdrive/fire_project/DS421_proj/output/prediction_fig.png", height=5, width=8, units="in", res = 100)
prediction_fig
dev.off()


```



To do 

results figures

correlation matrix
make top model table
make table of confusion matrix
figure showing class prob between true and false
figures of log odds for top two variables

fix figure from round 1 analysis

methods figures

study area map
1 map showing some vars and landcover outcomes (screen shot from gee)


update report
update slides for presentation

clean up state code (R and earth engine), put on github
make a journal entry about where things stand and what I need to do going foward to publish progress on this








