
---
title: "Rachel_Elif_analysisQ1.Rmd"
author: "DS421 Team"
date: "April 20, 2020"
output: md_document
---
Install Packages
```{r}
library(tidyverse)
library(foreign)
source('scripts/adams_theme.R') #for ggplot2
```

Import data
```{r}
df <- readRDS('data/cleanDataApril20_v2.Rds')
#this csv describe each ecotype code
eco_type_codes <- read_csv('data/landfire_ecotype_codes.csv')

landfire_codes <- foreign::read.dbf("data/US_200EVT\\US_200EVT.dbf")
landfire_codes <- landfire_codes %>% dplyr::select(VALUE,EVT_PHYS) %>% rename(landfire2016 = VALUE)

n_pixels <- nrow(df) #number of pixels in the dataset 
n_pixels
```


```{r}
#cleaning
df <- df %>%
  left_join(landfire_codes) %>%
  mutate(conifer = case_when(
    (EVT_PHYS == "Conifer" | EVT_PHYS == "Conifer-Hardwood") ~ TRUE,
    TRUE ~ FALSE
  ))
str(df)

#correcting temp

df$AvGTemp = df$AvGTemp / 12
str(df)
```

exporting fields to make field descriptions
```{r}
#names(df) %>% as.tibble() %>% write_csv(path = "data/data_fields.csv")
```


creating dataframe at the SUID level
```{r}
SUID_df <- df %>%
  mutate_at(.vars = "SUID", .funs = as.factor) %>%
  drop_na(SUID) %>%
  group_by(SUID) %>%
  summarise_if(.predicate = is.numeric, .funs = mean) 
```

looking at SUID_df
```{r}
str(df)
str(SUID_df)
SUID_df$latestSurveyNatDensity
```


Looking at distribution of response, latestSurveyNatDensity
```{r}
#looking at distributions...
hist(SUID_df$latestSurveyNatDensity, xlab = "Seedling Density", main = "Histogram of Natural Regeneration Seedling Desnity")


mean(SUID_df$latestSurveyNatDensity)
#151
var(SUID_df$latestSurveyNatDensity)
#118805

#variance >> mean, overdispered

hist(SUID_df$elevation)
plot(SUID_df$AvGppt, SUID_df$latestSurveyNatDensity)

```

Looking at pairs to note any obvious colinearity
```{r}
#pairs at the SUID level
with(SUID_df, pairs(SUID_df[, c('latestSurveyNatDensity', 'burnSev', 'slope', 'elevation', 'CWD', 'SolarLoad')])) #, 'SPB', 'SPO', 'SPMec', 'LAND_SUITA', 'PRODUCTIVI')]))

with(SUID_df, pairs(SUID_df[, c('latestSurveyNatDensity', 'burnSev', 'slope', 'elevation', 'CWD', 'AvGTemp', 'AvGppt')]))

#CWD is a function of temp and precip, correct? so should temp & precip not be included as features in the model as their variation should be captured by CWD??

with(SUID_df, pairs(SUID_df[, c('latestSurveyNatDensity', 'burnSev', 'LAND_SUITA', 'PRODUCTIVI', 'SPB')]))
```


Q1: What variables predict density of naturally regenerating seedlings (in an SUID) within 10 years of a fire? 

To answer this, we want to model latestSurveyNatDensity (in the dataframe SUID_df) as a function of: 
burnSev, slope, elevation, CWD, SolarLoad, temp, precip, etc.

Site preparation variables include: SPB, SPO, and SPMec, plantedFirst5Years, PlantedAfter5Yrs


```{r}
library(gamsel)

# GAM with feature selection
#this package also allows you to do cross validation
#notes: northness = cos(aspect), eastness = sin(aspect)
#CWD = some function of precip & temp & solar load ?


#the SUID_df only includes 360 observations
#for this gam, using all pixels where there's a recorded latestSurveyNatDensity
filtered_df = df[!is.na(df$latestSurveyNatDensity),]
str(filtered_df)
hist(filtered_df$latestSurveyNatDensity)
```


```{r}
#including all variables for gamsel
x <- as.matrix(filtered_df %>% dplyr::select(c(burnSev, forestType, SAP_surv_conifers, SAP_burn, elevation, slope, northness, eastness, CWD, AvGppt, AvGTemp, plantedFirst5Years, plantedAfter5Yrs, LAND_SUITA, PRODUCTIVI, SPB, SPO, SPMec)))
y <- as.numeric(filtered_df$latestSurveyNatDensity)

#standardizing so that gamsel can fit
#scale will do this, scale = F sets col-wise standardization 
x_standardized = scale(x, scale = FALSE)

hist(x_standardized[,5])

#For gamsel package degrees = basis functions. Burn sev, forest type, SPO, SPB, SPMech are binary or categorical, which is why I'm setting their "degree" to 1
#gamsel will choose lamda, i.e. the penalty term 
fit_gamsel<-gamsel(x_standardized, y, degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1,1, 1, 1, 1))
summary(fit_gamsel)
plot(fit_gamsel, newx = x_standardized)

#LAND_SUITA negative relationship
#PRODUCTIVI positive relationship --> what goes into these two indices? I'm surprised they have opposite effects
```


```{r}
#LOG TRANSFORM
y_logt <- log(1 + y)

#for reference: 
#x <- as.matrix(filtered_df %>% dplyr::select(c(burnSev, forestType, SAP_surv_conifers, SAP_burn, elevation, slope, northness, eastness, CWD, AvGppt, AvGTemp, plantedFirst5Years, plantedAfter5Yrs, LAND_SUITA, PRODUCTIVI, SPB, SPO, SPMec)))
fit_gamsel_logt<-gamsel(x_standardized, y_logt, degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1,1, 1, 1, 1))

summary(fit_gamsel_logt)

plot(fit_gamsel_logt, newx = x_standardized)

#same general trends,see v4 SAP_burn and v10 AvGppt, but less non-linearity using the log-transformed natregendensity values

```

```{r}
#Looking more closely at log(y)
hist(y_logt, main = "Histogram of Log-Natural Regeneration Density")
#Bimodal distribution?
#Artifact of sampling strategy?
#One thing we COULD do here is try to first predict where the regen densitiy is not 0

#this contains pixels where regen density is not 0
y_filt = y[y_logt != 0]
x_filt = x_standardized[y_logt != 0,]
str(x_filt)
#looking at for how many pixels regen density = 0
y_logt == 0


#filtered, predicting pixels regen density where regen != 0
fit_gamsel_filtered<-gamsel(x_filt, y_filt, degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1,1, 1, 1, 1))
summary(fit_gamsel_filtered)

#filtered w log transformed y
#filtered gamsel =predicting where regen density is NOT 0, binomial family
fit_gamsel_logt_filt<-gamsel(x_standardized, y_logt != 0, num_lambda = 200,family="binomial", degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1,1, 1, 1, 1))
summary(fit_gamsel_logt_filt)


```

Cross validation using gamsel:

```{r}
#cross validation 
#If we were to use the simple mean of y_filt as our prediction for all pixels, the root mse would be...
#sqrt(var(y_filt))
#sqrt(var(predict.gamsel(fit_gamsel_logt, x_filt, 200) - y_filt))

#cv on original model, y w/o log transform
cv_y= cv.gamsel(x_standardized, y, lambda = c(2.0, 1.75, 1.5, 1.25, 1.0, 0.75, 0.5, 0.25, 0.1, 0.05, 0.01, 0.0), degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1, 1, 1, 1, 1), type.measure = "mse")

cv_y
#using decreasing #s of lambda to find smallest possible val of prediction error
# $cvm within cv object = prediction error (with a pred error value for each value of lambda)
min(cv_y$cvm)
sd_pred_error = sqrt(min(cv_y$cvm))
sdy = sd(y)
#root mean sq error you would get if you just predicted a constant (the mean)

#% decrease in pred error as compared to error using mean = 36.6%
#model is only 36% better than using the mean y as predictor
((sdy - sd_pred_error) / sdy) *100


cv_logt_y = cv.gamsel(x_standardized, y_logt, lambda = c(2.0, 1.75, 1.5, 1.25, 1.0, 0.75, 0.5, 0.25, 0.1, 0.05, 0.01, 0.0), degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1, 1, 1, 1, 1), type.measure = "mse")

cv_logt_y

#sd deviation of the error in predicting the log transformed y
sd_pred_error_log = sqrt(2.39)
sd_log_y = sd(y_logt)
sd_pred_error_log_multipilicative = exp(sqrt(2.39))
#multiplicative factor, prediction error
sd_pred_error_log_multipilicative
#using this model, predictions are within a factor pf ~5 of the actual regen density


```


```{r}
#started working on this question of modelling where regen is not 0, using a binomial distribution, but I don't think it makes sense to go this route without first looking into their strategy for siting sampling plots
#need to know more about the system 

hist(as.numeric(y_logt != 0))
sum(y_logt != 0)/length(y_logt)
hist(y)
sum(y == 0)/length(y)
1.0 - (sum(y_logt != 0)/length(y_logt))


cv_logistic = cv.gamsel(x_standardized, y_logt != 0, family="binomial", lambda = c(2.0, 1.5, 1.0, 0.5, 0.4, 0.3, 0.25, 0.125, 0.0), degrees=c(1, 1, 10, 10, 10,10,10,10,10, 10, 10, 1,1,1, 1, 1, 1, 1), type.measure = "class")

cv_logistic
```

Below are several gams (fit with and without management) and glms (fit with and without management):

```{r}
library(gam)

#now using SUID_df 
#quasipoisson

fit1_gam <- gam(formula = latestSurveyNatDensity ~ burnSev + slope + elevation + CWD +  northness + eastness, data = SUID_df, family = quasipoisson())
summary(fit1_gam)

fit2_gam <- gam(formula = latestSurveyNatDensity ~ burnSev + forestType + slope + elevation + AvGTemp + AvGppt+ CWD +  northness + eastness + SPB+ SPO + SPMec + plantedFirst5Years + plantedAfter5Yrs, data = SUID_df, family = quasipoisson())
summary(fit2_gam)

anova(fit1_gam, fit2_gam)


SUID_df

fit2.5_gam <- gam(formula = latestSurveyNatDensity ~ burnSev + forestType + slope + aspect + elevation + AvGTemp + AvGppt+ SolarLoad + landfire2016 + CWD + northness + eastness + SPB+ SPO + SPMec + plantedFirst5Years + plantedAfter5Yrs + LAND_SUITA + PRODUCTIVI, data = filtered_df, family = quasipoisson())

summary(fit2.5_gam)
```



```{r}
library(MASS)
#need the MASS package to use negative binomial
#negative biomial bc variance is much larger than mean, see wiki for details: https://en.wikipedia.org/wiki/Negative_binomial_distribution#Poisson_distribution

Fit1_nb <- glm.nb(formula = latestSurveyNatDensity ~ burnSev + forestType +slope + elevation + CWD+ northness + eastness, data = SUID_df)
summary(Fit1_nb)

Fit2_nb <-glm.nb(formula = latestSurveyNatDensity ~ burnSev + forestType +slope + elevation + CWD+ northness + eastness + SPB + SPO + SPMec + plantedFirst5Years + plantedAfter5Yrs, data = SUID_df)
summary(Fit2_nb)
#did not converge

anova(Fit1_nb, Fit2_nb)


#with filtered df: 
#removing LAND_SUITA
#removing planted -- correlated w site potential?


Fit2.5_nb <-glm.nb(formula = latestSurveyNatDensity ~ burnSev + slope + elevation + northness + eastness + AvGppt + AvGTemp + SPB + SPO + SPMec + PRODUCTIVI, data = filtered_df)

summary(Fit2.5_nb)

SUID_df$LAND_SUITA

```


```{r}
#GLM
#environmental + management + FS measures of productivity/suitability
#What does FS use to create suitability & productivity indices?

str(SUID_df)
Fit3 <- lm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn +slope + elevation + CWD+ SolarLoad + LAND_SUITA + PRODUCTIVI, data = SUID_df)
summary(Fit3)

#environmental + FS land suitabiltiy & productivity indices
Fit4 <- lm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn +slope + elevation + CWD+ SolarLoad + LAND_SUITA + PRODUCTIVI + SPB + SPO + SPMec, data = SUID_df)
summary(Fit4)

#productivity most predictive

anova(Fit3, Fit4)



```
No significan difference between models where productivity & suitability indices are included


```{r}
#all env variables
Fit7 <- lm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn + elevation + slope + aspect + northness + eastness + CWD + AvGppt + AvGTemp + SolarLoad, + PRODUCTIVI, data = SUID_df)
summary(Fit7)

Fit8 <- lm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn + elevation + slope + aspect + northness + eastness + CWD + AvGppt + AvGTemp + SolarLoad + PRODUCTIVI+ SPB + SPO + SPMec + plantedAfter5Yrs + plantedFirst5Years, data = SUID_df)
summary(Fit8)

plot(SUID_df$PRODUCTIVI, SUID_df$AvGppt)

```

```{r}
layout(matrix(c(1, 2, 3, 4), 2, 2))
plot(Fit1)
plot(Fit2)
```

```{r}
#fit 8 is a GLM with all vars included
#AIC for variable selection 
?step
Fit9 <- glm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn + elevation + slope + aspect + northness + eastness + CWD + AvGppt + AvGTemp + SolarLoad + PRODUCTIVI+ SPB + SPO + SPMec + plantedAfter5Yrs + plantedFirst5Years, data = filtered_df, family= quasipoisson())

summary(Fit9)



Fit10 <- lm(formula = latestSurveyNatDensity ~ burnSev + forestType + SAP_surv_conifers + SAP_burn + elevation + slope + aspect + northness + eastness + CWD + AvGppt + AvGTemp + SolarLoad + PRODUCTIVI+ SPB + SPO + SPMec + plantedAfter5Yrs + plantedFirst5Years, data = filtered_df)
summary(Fit8)
summary(Fit8)
summary(Fit9)
AICfit <- step(Fit10, trace = 0)
summary(AICfit)


Fit11 <- glm(formula = latestSurveyNatDensity ~  elevation + slope + northness + eastness + AvGppt + AvGTemp + PRODUCTIVI + SPB + SPO + SPMec, data = filtered_df, family= quasipoisson())

summary(Fit11)
```




```{r}
#run xtable on "summary(of your moodel) " --> latex code to make table 
#render latex code online
#screenshot
library(xtable)
xtable(summary(Fit11))
```

